---
title: "Test_ComPsy"
author: "Felix Pichardo"
date: "2024-08-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages
library(plyr)   #for revalue
library(dplyr)  #for %>%
library(ggplot2)
library(gridExtra)
library(grid)   #for textGrob
library(kableExtra) #for knitr tables
library(tidyr)
library(here)
library(foreign)
library(bayesplot)
library(posterior)
library(cmdstanr)

# Directories
PROJ_DIR            = file.path(here::here())
DATA_DIR            = file.path(PROJ_DIR, "Data")
SAFE_DATA_DIR       = file.path(DATA_DIR, "AHRB")
MODELS_DIR          = file.path(PROJ_DIR, "models")
MODELS_RDS_DIR      = file.path(MODELS_DIR, "rds")
MODELS_RDS_FIT_DIR  = file.path(MODELS_RDS_DIR, "fit")
MODELS_RDS_PoPC_DIR = file.path(MODELS_RDS_DIR, "postpc")


# Source helper functions
source(file.path(PROJ_DIR, "scripts", "helper_functions_cmdSR.R"))

# Load Data
## Trial-level data
wave1.sav.file = file.path(SAFE_DATA_DIR, "modigt_data_Wave1.sav")
wave1.raw      = read.spss(wave1.sav.file,to.data.frame = TRUE)
```


# EV: 4 Param
## Single
```{r}
model_str = "igt_mod_sing_new_pvl_ddm"

single.fit.model.path = file.path(MODELS_RDS_FIT_DIR, paste0(model_str, "_fit.rds"))
rds_pathname  = single.fit.model.path

#sample_sub = sample(unique(wave1.raw$sid), 1)
sample_sub = "202903"
#sample_sub = "201401" #sample sub 2
sample_sub = "202111"

sample_df = wave1.raw %>%
  filter(sid == sample_sub)
```

### Prep Data
```{r}
data_to_extract =  c("N", "T", "Tsubj", "RTbound", "minRT", "RT", "choice", "shown", "outcome")
model_params = c("boundary", "tau", "beta", "drift_con", "exp_upd", "lambda", "alpha", "A", "update_pe", "exp_max")

data_list = extract_sample_data(sample_df,data_to_extract)

stanmodel_arg = readRDS(single.fit.model.path)
```

### Fit
```{r}
debug_flag = F
testing = T

if ((grepl('debug', rds_pathname) > 0) | (debug_flag)){
  # If this is a debug script
  n_warmup = 10
  n_iter = 20 # How many iterations we ACTUALLY want - this will have n_warmup added to it!
  n_chains = 1
  set_adapt_delta = 0.9
  set_max_treedepth = 10
  parallel_bool = F
} else if (testing) {
  # Testing script
  n_warmup = 100
  n_iter = 1000 # How many iterations we ACTUALLY want - this will have n_warmup added to it!
  n_chains = 2
  set_adapt_delta = 0.9
  set_max_treedepth = 10
  parallel_bool = T
}else {
  n_warmup = 1000
  n_iter = 10000 # How many iterations we ACTUALLY want - this will have n_warmup added to it!
  n_chains = 2
  set_adapt_delta = 0.95
  set_max_treedepth = 12
  parallel_bool = T
}

paste0("Running sampling with ", n_chains, " chains of ", n_warmup, " warmup up iterations and ", n_iter, " sampling iterations (", n_warmup + n_iter, " actual iterations per chain).")
paste0("Adapt delta: ", set_adapt_delta)
paste0("Max Tree Depth: ", set_max_treedepth)
paste0("Running in parallel: ", parallel_bool)

fit_new_ev_ddm_sing = fit_stan_model(stanmodel_arg, data_list,
                          n_warmup = n_warmup, n_iter = n_iter + n_warmup,
                          n_chains = n_chains, adapt_delta = set_adapt_delta,
                          max_treedepth = set_max_treedepth, parallel = TRUE)

model_fit_obj = fit_new_ev_ddm_sing
```

#### Save fit
```{r}
fit_output_file = sub('/models/', '/data/', paste0(tools::file_path_sans_ext(rds_pathname), "_fit_output.rds"))

saveRDS(model_fit_obj$fit, file = fit_output_file)
```


#### Extract Parameters
```{r}
model_fit_obj$params = extract_params(model_fit_obj$all_params, main_params_vec = model_params)

model_fit_obj$params = unname(model_fit_obj$params)
```


####
#
####
##Hier Group
```{r}
model_str = "igt_mod_group_hier_ev"

group.hier.ev.model.fit.path = file.path(MODELS_RDS_FIT_DIR, paste0(model_str, "_fit.rds"))
rds_pathname  = group.hier.ev.model.fit.path

sample_df = wave1.raw
```

### Prep Data
```{r}
data_to_extract = c("N", "T", "Tsubj", "RTbound", "minRT", "RT", "choice", "shown", "outcome")
model_params = c("boundary", "tau", "beta", "drift_con", "exp_upd", "lambda", "alpha", "A", "update_pe", "exp_max")
```

### Load fit
```{r}
fit_output_file = sub('/models/', '/Data/', paste0(tools::file_path_sans_ext(rds_pathname), "_output.rds"))

model_fit_obj = readRDS(file = fit_output_file)
```


### Diagnostics
#### Divergences
```{r}
run_selected_diagnostics(model_fit_obj, steps_to_run = c("divergences"))
```

#### Traceplots
```{r}
run_selected_diagnostics(model_fit_obj, steps_to_run = c("traceplots"))
```

#### Display chain density plots (10 per page)
```{r}
run_selected_diagnostics(model_fit_obj, steps_to_run = c("density_plots_by_chain"), plots_pp = 10)
```

#### Display overall density plots (10 per page)
```{r}
run_selected_diagnostics(model_fit_obj, steps_to_run = c("overall_density_plots"), plots_pp = 10)
```

#### R-hat
```{r}
run_selected_diagnostics(model_fit_obj, steps_to_run = c("rhat_all"), params = model_fit_obj$all_params)
```

#### Effective Sample Size
```{r}
run_selected_diagnostics(model_fit_obj, steps_to_run = c("ess_all"), params = model_fit_obj$all_params, lower_than_q = 0.1)
```

#### Monte Carlo Standard Error
```{r}
run_selected_diagnostics(model_fit_obj, steps_to_run = c("mcse_all"), params = model_fit_obj$all_params)
```

#### Autocorrelation
```{r}
run_selected_diagnostics(model_fit_obj, steps_to_run = c("autocorrelation"), params = head(model_fit_obj$params, 8))
```

#### Parallel Coordinates Plot
```{r}
run_selected_diagnostics(model_fit_obj, steps_to_run = c("parallel_coordinates"), params = head(model_fit_obj$params, 12))
```

#### Pairs Plot
```{r}
run_selected_diagnostics(model_fit_obj, steps_to_run = c("pairs_plot"), params = head(model_fit_obj$params, 8))
```


# EV DDM: 4 Param
##Hier Group
```{r}
model_str = "igt_mod_group_hier_ev_ddm"

group.hier.ev.model.fit.path = file.path(MODELS_RDS_FIT_DIR, paste0(model_str, "_fit.rds"))
rds_pathname  = group.hier.ev.model.fit.path

sample_df = wave1.raw
```

### Prep Data
```{r}
data_to_extract = c("N", "T", "Tsubj", "RTbound", "minRT", "RT", "choice", "shown", "outcome")
model_params = c("boundary", "tau", "beta", "drift_con", "exp_upd", "lambda", "alpha", "A", "update_pe", "exp_max")
```

### Load fit
```{r}
fit_output_file = sub('/models/', '/Data/', paste0(tools::file_path_sans_ext(rds_pathname), "_output.rds"))

model_fit_obj = readRDS(file = fit_output_file)
```


### Diagnostics
#### Divergences
```{r}
run_selected_diagnostics(model_fit_obj, steps_to_run = c("divergences"))
```

#### Traceplots
```{r}
run_selected_diagnostics(model_fit_obj, steps_to_run = c("traceplots"))
```

#### Display chain density plots (10 per page)
```{r}
run_selected_diagnostics(model_fit_obj, steps_to_run = c("density_plots_by_chain"), plots_pp = 10)
```

#### Display overall density plots (10 per page)
```{r}
run_selected_diagnostics(model_fit_obj, steps_to_run = c("overall_density_plots"), plots_pp = 10)
```

#### R-hat
```{r}
run_selected_diagnostics(model_fit_obj, steps_to_run = c("rhat_all"), params = model_fit_obj$all_params)
```

#### Effective Sample Size
```{r}
run_selected_diagnostics(model_fit_obj, steps_to_run = c("ess_all"), params = model_fit_obj$all_params, lower_than_q = 0.1)
```

#### Monte Carlo Standard Error
```{r}
run_selected_diagnostics(model_fit_obj, steps_to_run = c("mcse_all"), params = model_fit_obj$all_params)
```

#### Autocorrelation
```{r}
run_selected_diagnostics(model_fit_obj, steps_to_run = c("autocorrelation"), params = head(model_fit_obj$params, 8))
```

#### Parallel Coordinates Plot
```{r}
run_selected_diagnostics(model_fit_obj, steps_to_run = c("parallel_coordinates"), params = head(model_fit_obj$params, 12))
```

#### Pairs Plot
```{r}
run_selected_diagnostics(model_fit_obj, steps_to_run = c("pairs_plot"), params = head(model_fit_obj$params, 8))
```




